---
title: "stats306_ch04"
author: "Sharon Ma"
date: "1/5/2023"
output: html_document
---

# Stats 306: Lab 01 & Ch 01-04 

## Lab 1: Installing Jupyter Notebook and R

OH: Thurs, 4-5:30 G219 AH 
kathyhuo$@$umich.edu

## R for Data Science - 1 Introduction

```{r cars}
install.packages("tidyverse")
install.packages(c("nycflights13","gapminder","Lahman"))
```

```{r pressure, echo=FALSE}
library(tidyverse)
library(nycflights13)
library(gapminder)
library(Lahman)
```

## R for Data Science - 2 Data Visualization

```{r}
install.packages("palmerpenguins")
```

```{r}
library(palmerpenguins)
view(penguins) # opens data in a new tab
glimpse(penguins) 
?penguins # this and the one belong both do the same thing
help(penguins)
```

```{r}
ggplot(data=penguins) # creates an empty graph
```

```{r}
ggplot(data=penguins,
       mapping=aes(x=flipper_length_mm,y=body_mass_g)) # maps a specific variable 
# onto the x axis, another variable onto the y axis
```

```{r}
ggplot(data=penguins,
       mapping=aes(x=flipper_length_mm,y=body_mass_g))+
  geom_point() # adds a layer of points to the plot 
```

We now want to incorporate the different species into our plot by separating the
respective points into different colors. This is an aesthetics things, so let's 
put this condition in the aesthetic mapping function.

```{r}
ggplot(data=penguins,
       mapping=aes(x=flipper_length_mm,y=body_mass_g,color=species))+
  geom_point() 
```

We now want to add a smooth curve displaying the relationship b/w body mass and 
flipper length into our plot. This is a new graph (a curve) using a new geom (smooth), 
so add it.

```{r}
ggplot(data=penguins,
       mapping=aes(x=flipper_length_mm,y=body_mass_g,color=species))+
  geom_point()+
  geom_smooth()
```

We now have curves, but we want only one curve for the entire dataset, as opposed 
to separate curves for each species, so move this aesthetic over to the geom smooth 
function. 

```{r}
ggplot(data=penguins,
       mapping=aes(x=flipper_length_mm,y=body_mass_g))+
  geom_point(mapping=aes(color=species))+
  geom_smooth()
```

For accessibility purposes, add shapes to the different points that represent 
different species. 

```{r}
ggplot(data=penguins,
       mapping=aes(x=flipper_length_mm,y=body_mass_g))+
  geom_point(mapping=aes(color=species,shape=species))+
  geom_smooth()
```

We can now add labels to our plot by using the labs function. 

```{r}
ggplot(data=penguins,
       aes(x=flipper_length_mm,y=body_mass_g))+
  geom_point(aes(color=species,shape=species))+
  geom_smooth()+
  labs(title="Body Mass and Flipper Length",
       x="Flipper Length (mm)",
       y="Body Mass (g)",
       color="Species",
       shape="Species")
```

Q: Make a scatterplot of bill_depth_mm vs bill_length_mm. Describe the relationship 
b/w these two variables.

A: There is no relationship b/w bill depth and length. Bill length flunctuates around
45mm on avg, regardless of bill depth. 

```{r}
ggplot(data=penguins,
       aes(x=bill_depth_mm,y=bill_length_mm))+
  geom_point()+
  geom_smooth()+
  labs(title="Bill Depth vs Length",
       x="Bill Depth (mm)",
       y="Bill Length (mm)")
```

Q: What happens if you make a scatterplot of species vs bill_depth_mm? Why is 
the plot not useful?

A: Not useful b/c not too diff than regular distribution of bill depth amongst 
all penguins.

```{r}
ggplot(data=penguins,
       aes(x=species,y=bill_depth_mm))+
  geom_point()
```

Q: Why does the following give an error and how would you fix it? 

A: Missing the aesthetics for x and y.

```{r}
# ggplot(data=penguins)+geom_point()
```

Q: What does the na.rm argument do in geom_point()? What is the default value of 
the argument? Create a scatterplot where you successfully use this argument set to 
TRUE. 

A: The default is na.rm=FALSE, so don't remove missings. 

```{r}
ggplot(data=penguins,
       aes(x=species,y=bill_length_mm))+
  geom_point()
```

```{r}
ggplot(data=penguins,
       aes(x=species,y=bill_length_mm))+
  geom_point(na.rm=TRUE)
```

```{r}
ggplot(data=penguins,
       aes(x=species,y=bill_length_mm))+
  geom_point(na.rm=FALSE)
```

```{r}
ggplot(penguins,aes(x=species))+geom_bar()
```

```{r}
ggplot(penguins,aes(x=body_mass_g))+geom_histogram(binwidth=200)
```

```{r}
ggplot(penguins,aes(x=body_mass_g))+geom_density()
```

Shows the distribution of body mass amongst penguins.

Q: Make a bar plot of species of penguins, where you assign species to the y aesthetic. How is this plot different?

A: Flips the axes of the graph around. 

```{r}
ggplot(penguins,aes(y=species))+geom_bar()
```

Q: How are the following two plots different? Which aesthetic, color or fill, is 
more useful for changing the color of the bars?

A: Simply outlines the bars vs fills the bars. The ladder is more useful. 

```{r}
ggplot(penguins,aes(x=species))+geom_bar(color="red")
```

```{r}
ggplot(penguins,aes(x=species))+geom_bar(fill="red")
```

```{r}
ggplot(penguins,aes(x=species,y=body_mass_g))+geom_boxplot()
```

```{r}
ggplot(penguins,aes(x=island,fill=species))+geom_bar()
```

```{r}
ggplot(penguins,aes(x=island,fill=species))+geom_bar(position="fill")
```

IMPORTANT: Changes to proportions! 

```{r}
ggsave("insertname.png") # saves the most recent plot
```



